Fail Faster

### 2025-03-12 paper/paper.tex
blah

The power of staging for optimization away abstraction overheads comes from
defining functions that accept and return \texttt{code} values.  A function
\texttt{f : 'a code -> 'b code} is a function that takes a program computing a
run-time \texttt{'a} and transforms it into a program computing a run-time
\texttt{'b}. In particular, because \texttt{f} itself runs at compile time, the
abstraction of using \texttt{f} is necessarily completely eliminated by
run-time.
A code-transforming function \texttt{'a code -> 'b code} can also be converted \emph{code for a function} --- a value of type
\texttt{('a -> 'b) code} --- with the following program:
\begin{lstlisting}
let eta (f : 'a code -> 'b code) : ('a -> 'b) code = .<fun x -> .\~(f .<x>.)>.
\end{lstlisting}

@@

The crux of our binding time analysis is that particular size random seed are
only known at run-time (the later stage), but the code of the generator itself
is known at compile time.  This means our library's generator type \texttt{'a
Gen.t} should be \texttt{int code -> SR.t code -> 'a code}: a compile-time
function from dynamically-known size and seed to dynamically-determined result.

@@

\jwc{This is 3-stage metaprogramming! There's PPX-time, compile time, and run-time.}

>>>>

The power of staging for optimization away abstraction overheads comes from
defining functions that accept and return \texttt{code} values.  A function
\texttt{f : 'a code -> 'b code} is a function that takes a program computing a
run-time \texttt{'a} and transforms it into a program computing a run-time
\texttt{'b}. In particular, because \texttt{f} itself runs at compile time, the
abstraction of using \texttt{f} is necessarily completely eliminated by
run time.
A code-transforming function \texttt{'a code -> 'b code} can also be converted \emph{code for a function} --- a value of type
\texttt{('a -> 'b) code} --- with the following program:
\begin{lstlisting}
let eta (f : 'a code -> 'b code) : ('a -> 'b) code = .<fun x -> .\~(f .<x>.)>.
\end{lstlisting}

@@

The crux of our binding time analysis is that particular size random seed are
only known at run time (the later stage), but the code of the generator itself
is known at compile time.

Generators --- values of type \texttt{'a Bq.t} --- are never constructed dynamically at run time in practice,
so the

This means our library's generator type \texttt{'a
Gen.t} should be \texttt{int code -> SR.t code -> 'a code}: a compile-time
function from dynamically-known size and seed to dynamically-determined result.

@@

\jwc{This is 3-stage metaprogramming! There's PPX time, compile time, and run time.}
